Digit Recognition: Nearest Neighbors vs. Perceptron

### 2013-06-09 Paper/paper.tex
Update writeup code

\section{The Challenge}

Optical Character Recognition (OCR) is a computational problem where a machine
attempts to classify an image of text. We tackled a subset of this problem,
attempting to classify the digits 0-9 from data provided by the Kaggle Digit
Recognition
competition\footnote{\url{http://www.kaggle.com/c/digit-recognizer}}.

The Kaggle data provides two groups of data, 42,000 training samples with labels
and 28,000 unlabeled testing samples.

\begin{table}[ht]
\caption{Subset of Kaggle Data} % title
\centering 
\begin{tabular}{c c c c }

Label & Pixel 77 & Pixel 78 & Pixel 79 \\

8 & 0 & 0 & 0 \\
6 & 89 & 208 & 135 \\ [1ex]

\end{tabular}
\label{table:nonlin}
\end{table}

\section{General formatting instructions}
\label{gen_inst}

The text must be confined within a rectangle 5.5~inches (33~picas) wide and
9~inches (54~picas) long. The left margin is 1.5~inch (9~picas).
Use 10~point type with a vertical spacing of 11~points. Times New Roman is the
preferred typeface throughout. Paragraphs are separated by 1/2~line space,
with no indentation.

Paper title is 17~point, initial caps/lower case, bold, centered between
2~horizontal rules. Top rule is 4~points thick and bottom rule is 1~point
thick. Allow 1/4~inch space above and below title to rules. All pages should
start at 1~inch (6~picas) from the top of the page.

%The version of the paper submitted for review should have ``Anonymous Author(s)'' as the author of the paper.

For the final version, authors' names are
set in boldface, and each name is centered above the corresponding
address. The lead author's name is to be listed first (left-most), and
the co-authors' names (if different address) are set to follow. If
there is only one co-author, list both author and co-author side by side.

Please pay special attention to the instructions in section \ref{others}
regarding figures, tables, acknowledgments, and references.

\section{Headings: first level}
\label{headings}

First level headings are lower case (except for first word and proper nouns),
flush left, bold and in point size 12. One line space before the first level
heading and 1/2~line space after the first level heading.

\subsection{Headings: second level}

Second level headings are lower case (except for first word and proper nouns),
flush left, bold and in point size 10. One line space before the second level
heading and 1/2~line space after the second level heading.

\subsubsection{Headings: third level}

Third level headings are lower case (except for first word and proper nouns),
flush left, bold and in point size 10. One line space before the third level
heading and 1/2~line space after the third level heading.

>>>>

\section{The Problem}

\subsection{Image Data}
Optical Character Recognition (OCR) is a computational problem where a machine
attempts to classify an image of text. We tackled a subset of this problem,
attempting to classify the digits 0-9 from data provided by the Kaggle Digit
Recognition
competition\footnote{\url{http://www.kaggle.com/c/digit-recognizer}}.

The Kaggle data provides two groups of data, 42,000 training samples with labels
and 28,000 unlabeled testing samples. Each training sample contains a 48 pixel
by 48 pixel image containing a handwritten number. The image is represented as
an array of 768 pixel values which represent how dark each pixel of the image
is. Table 1 is an example of a training sample.

\begin{table}[ht]
\textbf{\caption{Subset of Kaggle Data}} % title
\centering 
\begin{tabular}{c | c c c }
Label & Pixel 77 & Pixel 78 & Pixel 79 \\
\hline
8 & 0 & 0 & 0 \\
6 & 89 & 208 & 135 \\ [1ex]
\end{tabular}
\label{table:nonlin}
\end{table}

\section{K Nearest Neighbors (KNN)}

\subsection{Overview}

K Nearest Neighbors is a relatively simple algorithm which often gets surprisingly
good results given its simplicity. Given a testing sample $X$, the algorithm
looks through all the training samples it has previously seen and finds the
\textit{K} most similar samples. The algorithm then implements some form of
voting mechanism among the \textit{K} samples to classify the testing sample as
a digit 0-9. The voting mechanism can either be as straightforward as selecting the
majority label, to something more complicated which involves weighting the
samples based on some distance metric.

\subsection{Implementation Details}

The first step in implementing KNN is to find the \textit{K} nearest
neighbors. This is accomplished by taking the euclidean distance between the
testing sample we are classifying and every training sample in our training data
set.

\begin{center}
$Euclidean(TestX, TrainX) = \sum\limits_{i \in TrainX} {(TestX - TrainX_i)^2}$
\end{center}

Because the image data is defined as an array of 768 pixels, $TestX - TrainX_i$
is actually doing a 768 dimension comparison between data points.

Once we identified the \textit{K} nearest neighbors to our testing sample, we
had to use this information to classify the testing point. We considered two
approaches when deciding how to perform this classification.

\subsubsection{Unweighted KNN}

The first method we implemented to classify our testing sample was a simple
majority voting algorithm which selects the majority label from the \textit{K}
closest neighbors. For example, if we ran the nearest neighbors algorithm with
$K = 5$ and 3 of the neighbors had a label of 6, we would classify our testing
sample as a 6.

\subsubsection{Weighted KNN}

Given the \textit{K} nearest neighbors, the weighted version of the algorithm
multiplies each label by a weight which indicates how close it is to our testing
sample. These weights help to emphasize points that are closer to our testing
sample while reducing the impact further away training points have on
classifying our testing sample.

\subsection{Cross Validation}

Our implementation of KNN has a number of parameters which require
tuning. In particular, we need to decide whether or not to use the weighted
version of our KNN algorithm and how many neighbors to consider when
implementing our algorithm.

To do this, we implemented K folds cross validation using 5,000 training
samples. The reduced training set allowed us to run our algorithm using various
configurations to see which configuration was most accurate.

\begin{table}[ht]
\textbf{\caption{KNN Cross Validation Error Rates}} % title
\centering
\begin{tabular}{c | c c }
K & Unweighted KNN Error & Weighted KNN Error \\
\hline
1 & .077 & .077 \\
2 & .092 & .092 \\
3 & .075 & .075 \\
4 & .078 & .073 \\
5 & .078 & .075 \\
10 & .085 & .079 \\
15 & .095 & .089 \\
20 & .102 & .098 \\ [1ex]
\end{tabular}
\label{table:nonlin}
\end{table}

Our cross validation results suggest that running weighted KNN with $K = 4$
yields the most successful classification rate. One caveat to this is that our
cross validation code ran using 5,000 training samples instead of the 42,000
samples in our full training set. It is possible that data we used to cross
validate is no representative of the full data set.

\subsection{Runtime}

\subsection{Results}

\section{Kernelized Perceptron}
\subsubsection{Overview}

\subsubsection{Implementation Details}

\subsubsection{Results}
